---
title: "Simple document"
output: html_document
---


```{r, echo=FALSE, message=FALSE}
library(tidyverse)
library(janitor)
library(readxl)
library(haven)
```
## Import some data

l want to import `FAS_litters.csv`,

```{r}
litters_df <- read.csv("data/FAS_litters.csv")
```

Yay! l want better names.

``` {r}
names(litters_df)
litters_df <- janitor::clean_names(litters_df)
```

```{r,eval=FALSE}
views(litters_df)
```

```{r}
skimr::skim(litters_df)
```

## Arguments in `read_csv`

```{r}
litters_df1 <- 
  read_csv(
    "data/FAS_litters.csv",
    skip = 5,
    col_names = FALSE ,
    na  = "Low8")
```
## Reading from Excel

Reading MLB data
```{r}
mlb11_df <- read_excel("data/mlb11.xlsx")
```
LotR words is next
```{r}
fellow_df <- read_excel("data/LotR_Words.xlsx",range="B3:D6")
```

##Read a SAS file
```{r}
pulse_df <- read_sas("data/public_pulse_data.sas7bdat",encoding="latin1")
```

why note use read.csv

`read.csv` doesn't give `tibble` and that's really bad
```{r,eval=FALSE}
litters_df2 <- read.csv("data/FAS_litters.csv")
litters_df2 <- janitor::clean_names(litters_df2)

pups_df <- read.csv("data/FAS_pups.csv")
pups_df <- janitor::clean_names(pups_df)
```

## how do l export data

there's a good way!
```{r}
write_csv(fellow_df,"data/fellowship_words.csv")
```

# lecture 2
## 'select'

Let's slect some columns

```{r}
select(litters_df2,group,litter_number)
##Choose the beginning columns
select(litters_df2,group,starts_with("pups"))
## you can use - to get rid of some columns
select(litters_df2,-litter_number)
## use GROUP to substitute group, and choose them
select(litters_df2,GROUP = group)
## rename to rename and only rename
rename(litters_df2,GROUP = group)
## relocate exchange two sites
relocate(litters_df2,)
```

## `filter`
let's get rid of rows
```{r}
filter(litters_df2,gd_of_birth == 20)
filter(litters_df2,group == "Con7")
filter(litters_df2,gd0_weight < 23)
##use ! for something that is not
filter(litters_df2,pups_survive !=4)
filter(litters_df2,!(group == "Con7"))
## use %in% set condition
filter(litters_df2,group %in% c("Con7","Con8"))
## use | for meeting only one condition
filter(litters_df2,group == "Con7"| gd_of_birth ==20)
##get rid of na
drop_na

```

## `mutate`
Let's add or change columns
```{r}
## add a new columb by two existing columns
## str_to_lower change the upper to lower
mutate(litters_df2,weight_change = gd18_weight - gd0_weight,group = str_to_lower(group))
```

## `arrange`
```{r}
arrange(litters_df2,gd_of_birth,gd0_weight)
```

## Pipes
```{r}
## tradtion way
litters_data_raw <-  read_csv("data/FAS_litters.csv")
litters_clean_name <- janitor::clean_names(litters_data_raw)
litters_select <- select(litters_clean_name,group,pups_survive)
litters_filtered <- filter(litters_select, group =="Con7")
##PIPE
litters_df <- 
  read.csv("data/FAS_litters.csv") %>%
  janitor::clean_names() %>%
  select(group, pups_survive) %>%
  filter(group ="Con7")

litters_df <- 
  read.csv("data/FAS_litters.csv") %>%
  janitor::clean_names() %>%
  select(-pups_survive) %>% 
  mutate(
    weight_change = gd18_weight -gd0_weight,
    group = str_to_lower(group)
  ) %>% 
  drop_na(weight_change) %>% 
  filter(group %in% c("con7","con8")) %>% 
  select(litter_number,group,weight_change,everything())
```

# lecture 3 tidy

## pivot longer
load the pulse data
```{r}
pulse_df <- 
  read_sas("data/public_pulse_data.sas7bdat")
  janitor::clean_names()

##将BDI的四列数据并到一列，用visit作为列名  
pulse_tidy <- 
  pulse_df %>% 
  pivot_longer(
    BDIScore_BL:BDIScore_12m,
    names_to="visit",
    names_prefix = "bdi_score_",
    values_to ="bdi"
  ) %>% 
  ##去除visit列内相同前缀
  mutate(
    visit = replace(visit, visit == "bl", "00m"),
    visit = factor(visit)
    ) 
  
```
pivot_longer over gather()

##pivot_wider
```{r}
analysis_df <- 
  tibble(
    group = c("treatment","treatment","control","control"),
    time = c("a","b","a","b"),
    group_mean =c (4,8,3,6)
  )

analysis_df %>% 
  pivot_wider(
    names_from = "time", 
    values_from = "group_mean"
  ) %>% 
  knitr::kable()
```

`pivot_wider()` over `spread()`
## bind_rows
import the LotR movie word stuff
```{r}
##import data from movie file
fellowship_df <- 
  read_excel("data/LotR_Words.xlsx",range="B3:D6") %>% 
  mutate(movie = "fellowship_rings")

two_towers_df <- 
  read_excel("data/LotR_Words.xlsx",range="F3:H6") %>% 
  mutate(movie = "two_towers")

return_df <- 
  read_excel("data/LotR_Words.xlsx",range="J3:L6") %>% 
  mutate(movie = "return_king")

##bind them together and pivot by sex
lotr_df <- 
  bind_rows(fellowship_df,two_towers_df,return_df) %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    female:male,
    names_to = "sex",
    values_to = "words"
  ) %>% 
  relocate(movie)
```
## joins

`bind_rows()` over `rbind`

Look at FAS data

```{r}
litters_df <- 
  read_csv("data/FAS_litters.csv") %>% 
  janitor::clean_names() %>% 
  ## the group include two variables
  separate(group,into=c("dose","day_of_tx"),3) %>% 
  ## let the key number be the first
  relocate(litter_number)
  mutate(dose = str_to_lower(dose))

pups_df <- 
  read.csv("data/FAS_pups.csv") %>% 
  janitor::clean_names() %>% 
  mutate(sex = recode(sex,`1`="male", `2`="female"))
```

Let's join these up
```{r}
fas_df <- 
  left_join(pups_df,litters_df,by="litter_number") %>% 
  relocate(litter_number,dose,day_of_tx)
```



